---
title: "Estimating the degree to which the miRNA expression profiles resemble either BRAFV600Eor RAS-mutated PTC with the BRAFV600E – RAS score"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load BRAFV600E Mutation VS RAS Mutation DE miRNA data
```{r,message=FALSE,warning=FALSE}
library(DESeq2)
load("miRNA_BRS_DE.Rdata")
rownames(res_fc1.5q0.05[1:min(nrow(res_fc1.5q0.05),100),])
```

# Get the centroid of BRAFV600E M and RAS M data
```{r,message=FALSE,warning=FALSE}
CB <- matrix(rowMeans(top100_vst_braf),ncol=1, byrow=T)
colnames(CB) <- c("centroid")
rownames(CB) <- rownames(top100_vst_braf)

CR <- matrix(rowMeans(top100_vst_ras),ncol=1, byrow=T)
colnames(CR) <- c("centroid")
rownames(CR) <- rownames(top100_vst_ras)
```

# Get z-score normalized Euclidean distance
```{r,message=FALSE,warning=FALSE}
library(distances)
TB <- lapply(seq(1,ncol(top100_vst_tu),by=1),function(i){distances(rbind(t(top100_vst_tu[,i]),t(CB)))})
TB <- sapply(TB, function(i){i[1,2]}) 
TB <- (TB-mean(TB))/sd(TB)

TR <- lapply(seq(1,ncol(top100_vst_tu),by=1),function(i){distances(rbind(t(top100_vst_tu[,i]),t(CR)))})
TR <- sapply(TR, function(i){i[1,2]}) 
TR <- (TR-mean(TR))/sd(TR)
```

# Generate BRAFV600E – RAS score according to the method described in the TCGA analysis

Briefly, we computed the centroids of the top Mut.BRAFV600E vs Mut.RAS DE miRNA signature in B, c(B), and R, c(R), and for each tumor t we defined the BRAFV600E-RAS score (BRS) of t as the difference between the distance of t from c(B) and c(R):
BRS(t) = |v(t)-c(B)|2 - |v(t)-c(R)|2
where v(t) is the vector of miRNA vst expression of the top DE miRNAs in t, and the distance |x-y|2 is the normalized Euclidean distance between the expression vectors.
Tumors with negative BRS (closer to B than to R) were defined BRAFV600E-like, while tumors with positive BRS (closer to R than to B) were defined RAS-like.
```{r,message=FALSE,warning=FALSE}
BRS <- TB-TR
BRS <- matrix(BRS,ncol=1, byrow=T)
colnames(BRS) <- c("BRS")
rownames(BRS) <- colnames(top100_vst_tu)  
write.csv(BRS,"tumor_BRS_miRNA.csv")
BRS
```
