---
title: "Generating the ERK-activity score of 52 expressed genes responsive to MEK inhibition"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
        pandoc_args: [
         "+RTS", "-k16g",
         "-RTS"
         ]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read in vst expression data of all samples
```{r,message=FALSE,warning=FALSE}
vst <- read.csv("totalRNA_vst.csv",header=T,row.names=1)
library(rmarkdown)
paged_table(vst)
```

# Read in sample table and select tumor samples
```{r,message=FALSE,warning=FALSE}
sample <- read.csv("totalRNA_sample_table.csv",header=T,row.names=1)
sample_tu <- subset(sample,Specimen.Type.ID == "Primary Tumor")
paged_table(sample_tu)
vst_tu <- vst[,colnames(vst)%in%rownames(sample_tu)]
```

# Read in the BRS score and seperate B-like samples and R-like samples 
```{r,message=FALSE,warning=FALSE}
BRS <- read.csv("tumor_BRS_totalRNA.csv",header=T,row.names=1)
B <- rownames(subset(BRS,BRS<0))
R <- rownames(subset(BRS,BRS>0))
```

# Get the vst expression of 52 ERK-activity related genes in tumor samples 
```{r,message=FALSE,warning=FALSE}
genes <- read.csv("52genes.csv",header=F)
genes_vst_tu <-vst_tu[rownames(vst_tu)%in% genes $ V1,]
write.csv(genes_vst_tu,"52genes_vst_inTumor.csv")
genes_vst_tu_B <- genes_vst_tu[,colnames(genes_vst_tu)%in%B]
genes_vst_tu_R <- genes_vst_tu[,colnames(genes_vst_tu)%in%R]
```

# Generate the ERK activity score according the the method described in the TCGA analysis

## Compute the mean and standard deviation of vst expression of the 52 genes in BRAFV600E-like and RAS-like tumor samples separately.
```{r,message=FALSE,warning=FALSE}
MB <- matrix(rowMeans(genes_vst_tu_B),ncol=1, byrow=T)
colnames(MB) <- c("mean_in_B")
rownames(MB) <- rownames(genes_vst_tu_B)
library(matrixStats)
SB <- matrix(rowSds(as.matrix(genes_vst_tu_B)),ncol=1, byrow=T)
colnames(SB) <- c("SD_in_B")
rownames(SB) <- rownames(genes_vst_tu_B)

MR <- matrix(rowMeans(genes_vst_tu_R),ncol=1, byrow=T)
rownames(MR) <- rownames(genes_vst_tu_R)
colnames(MR) <- c("mean_in_R")
SR <- matrix(rowSds(as.matrix(genes_vst_tu_R)),ncol=1, byrow=T)
colnames(SR) <- c("SD_in_R")
rownames(SR) <- rownames(genes_vst_tu_R)
```
## Compute the pooled mean and standard deviation
```{r,message=FALSE,warning=FALSE}
MP <- (MB* ncol(genes_vst_tu_B) + MR * ncol(genes_vst_tu_R)) / (ncol(genes_vst_tu_B) + ncol(genes_vst_tu_R))
colnames(MP) <- c("pooled_mean")
SP <- ( ((ncol(genes_vst_tu_B) - 1) * (SB^2) + (ncol(genes_vst_tu_R) - 1) * (SR^2)) / (ncol(genes_vst_tu_B) + ncol(genes_vst_tu_R) -2) )^(1/2)
colnames(SP) <- c("pooled_SD")
rownames(SP) <- rownames(genes_vst_tu)
```
## Compute the Z-score for each sample by subtracting the pooled mean from the vst expression and dividing by the pooled standard deviation
```{r, message=FALSE,warning=FALSE}
zs <- (genes_vst_tu - MP) / SP
```
## Generate th ERK-activity score for each sample:  the sum of the Z-scores for the genes in set A, minus the sum of Z-scores for the genes in set B
```{r,message=FALSE,warning=FALSE}
gene_A <- subset(genes, V2=="A")
gene_B <- subset(genes, V2=="B")
zs_A <- zs[rownames(zs)%in%(gene_A $ V1),]
zs_B <- zs[rownames(zs)%in%(gene_B $ V1),]
ERK <- colSums(zs_A)-colSums(zs_B)
ERK <- as.data.frame(ERK)
colnames(ERK) <- c("ERK_score")
sample_tu <- sample_tu[order(rownames(sample_tu)),]
ERK <- cbind(ERK,sample_tu[,c("FinalDriverPathway.10Jan2020","FinalDriverGroup.10Jan2020","FinalDriver.10Jan2020","TSS.Subject.ID","Dose.MGY","Age_exposure","Age_S")])
ERK <- cbind(ERK,BRS)
names(ERK)[ncol(ERK)]<- "BRS"
write.csv(ERK,"tumor_ERK_BRS")
paged_table(ERK)

```